### 2024-04-16 Cover Letter/main.tex
update for JAFR submission

\documentclass{article}

\usepackage{charter} 
\usepackage{ragged2e}
\usepackage{graphicx}
\usepackage[
	a4paper,    % Paper size
	top=1in,    % Top margin
	bottom=1in, % Bottom margin
	left=1in,   % Left margin
	right=1in,  % Right margin
	%showframe  % Uncomment to show frames around the margins for debugging purposes
]{geometry}

\setlength{\parindent}{0pt}     % Paragraph indentation 
\setlength{\parskip}{1em}       % Vertical space between paragraphs

\usepackage{graphicx}       % Required for including images

\usepackage{fancyhdr}       % Required for customizing headers and footers

\fancypagestyle{firstpage}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{0pt} % No header rule
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\fancypagestyle{subsequentpages}{%
	\fancyhf{} % Clear default headers/footers
	\renewcommand{\headrulewidth}{1pt} % Header rule thickness
	\renewcommand{\footrulewidth}{1pt} % Footer rule thickness
}

\AtBeginDocument{\thispagestyle{firstpage}} % Use the first page headers/footers style on the first page
\pagestyle{subsequentpages} % Use the subsequent pages headers/footers style on subsequent pages

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	FIRST PAGE HEADER
%----------------------------------------------------------------------------------------

\includegraphics[scale=0.5]{logo.jpg} % Logo

\vspace{-1em} % Pull the rule closer to the logo

\rule{\linewidth}{1pt} % Horizontal rule

\bigskip\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	YOUR NAME AND CONTACT INFORMATION
%----------------------------------------------------------------------------------------

\hfill
\begin{tabular}{l @{}}
	$16^{th}$ April, 2024\\ % Date
	\textbf{Bryce Ronald William Boyd $|$ PhD Candidate}\\
	Queensland University of Technology \\
	School of Mathematical Sciences\\~\\
	Address: Level 8 Y Block, 2 George Street, \\
	Brisbane, Queensland, Australia, 4000\\
	Phone: +61 400 310 105 \\
	Email: bryce.polley@awri.com.au
\end{tabular}

\bigskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	ADDRESSEE AND GREETING
%----------------------------------------------------------------------------------------

\begin{tabular}{@{} l}
	\textbf{Dr. John P. Brooks and Professor Yangchao Luo,}\\ Editors-in-Chief \\ Journal of Agriculture and Food Research
\end{tabular}

\bigskip % Vertical whitespace

Dear Editors,
\\~\\
\parskip % Vertical whitespace

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\centering \textbf{Submission of Manuscript for Consideration by \\Journal of Agriculture and Food Research}\\
\bigskip % Vertical whitespace
\justifying
\noindent

\bigskip \noindent% Vertical whitespace
We wish to submit a new manuscript entitled  ''\textit{An analysis of underlying relationships between factors related to operating costs and revenue in Australian vineyards}'' for your consideration by the Journal of Agriculture and Food Research.
\newline \noindent
In this manuscript we explore the contributing factors to Australian vineyards' operational costs and revenue. We provide insights into the interrelatedness of the factors contributing to operational cost and revenue, we further explore the relationships between these factors and regional differences; showing how different regional effects impact on vineyard operations and outcomes. We use a novel, extensive nationwide data from Australia, collected over a decade by Sustainable Winegrowing Australia. We construct and analyse several models of the multifaceted, multi-resource nature of vineyard management.
\newline \noindent
We believe that this manuscript is appropriate for publication by the Journal of Agriculture and Food Research as it demonstrates the regional differences in operational costs and revenue of Australian vineyards and further link these outcomes to contributing factors to rank the relative importance of different vineyard activities and resource use. We illustrate the economic outcomes of Australian vineyards and link these outcomes to different regional factors. Thank you for your consideration of this manuscript.
\bigskip 
\newline \noindent% Vertical whitespace
\includegraphics[scale=0.1]{signature_block.png}
\newline \noindent
Bryce Boyd (corresponding author) on behalf of all coauthors: Kate Helmstedt, Mardi Longbottom, Kerrie Mengersen\\
\newline \noindent
\textit{Corresponding Author}, \textit{Queensland University of Technology}

\end{document}